---
title: "p8105_hw3_scs2267"
output: github_document
date: "2024-10-16"
---

# Problem 1
```{r}
library(p8105.datasets)
data("ny_noaa")

library(tidyverse)
```

```{r}
weather_df = ny_noaa
```
This data frame contains information about precipitation, snowfall, maximum temperature, and minimum temperature by date.

(Unfortunately I was unable to complete the rest of this section)

# Problem 2

```{r}
demographic_df = read_csv("./data/demographic.csv", show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  mutate(
     sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"),
    sex = as.factor(sex))
  

accel_df = read_csv("./data/nhanes_accel.csv", show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    min1:min1440,
    names_to = "MIMs",
    values_to = "time"
  )

mims_df = left_join(accel_df, demographic_df, by = "seqn") %>% 
  drop_na() %>% 
  filter(age>=21)
```

## MIMs by Sex for Each Education Category

### Table
```{r}
sex_and_educ_df = group_by(mims_df, education, sex) %>% 
  summarize(observations = n())

sex_and_educ_df
```

The highest frequency in this table are males and females in education level 3. In education level 1, there appears to be a higher frequency of females than males, and in education level 2, there appears to be a higher frequency of males to females.

### Plot 1
```{r}
mims_df %>% 
  group_by(seqn, sex, age, time, education) %>% 
  summarise(total_activity = sum(time)) %>% 
  ggplot(aes(x = age, y = total_activity, color = sex))+
  facet_wrap(sex~education)+
  geom_point(alpha = .5)+
  geom_smooth(se = FALSE)+
  theme_minimal()
  
```
This plot indicates higher activity among young men and women in education level three than the other education groups. Men and women appear to have a similar distribution in the first two education groups, but men appear to have higher variability in activity when they are younger in the third education category.


### Plot 2
```{r}
mims_df %>% 
  group_by(sex, MIMs,time, education) %>% 
  ggplot(aes(x = MIMs, y = time, color = sex))+
  facet_wrap(~education)+
  geom_point(alpha = .5)
```
This table indicates that, among all education and sex groups, there appears to one peak in activity earlier in the day, and another peak in activity later in the day. Activity per minute appears to be equally distributed between men and women in all education groups, however, in education group 1, there appears to be higher variation in activity in men towards the beginning of the day, and higher variation in activity in women towards the end of the day. 


# Problem 3

```{r}
jan_2020_df = read_csv("./citibike/Jan 2020 Citi.csv.zip", show_col_types = FALSE) %>% 
  mutate(
    month = "January",
    year = "2020",
    month = as.factor(month),
    year = as.factor(year)
  )

july_2020_df = read_csv("./citibike/July 2020 Citi.csv.zip", show_col_types = FALSE) %>% 
  mutate(
    month = "July",
    year = "2020",
    month = as.factor(month),
    year = as.factor(year)
  )

jan_2024_df = read_csv("./citibike/Jan 2024 Citi.csv.zip", show_col_types = FALSE) %>% 
  mutate(
    month = "January",
    year = "2024",
    month = as.factor(month),
    year = as.factor(year)
    )

july_2024_df = read_csv("./citibike/July 2024 Citi.csv.zip", show_col_types = FALSE) %>% 
  mutate(
    month = "July",
    year = "2024",
    month = as.factor(month),
    year = as.factor(year)
    )

citibike_df = bind_rows(jan_2020_df, july_2020_df, jan_2024_df, july_2024_df) %>% 
  relocate(ride_id, year, month, member_casual) %>% 
  arrange(year, month)
```

## Table 1 
```{r}
member_table_df = group_by(citibike_df, year, month, member_casual) %>% 
  summarize(number_of_rides = n()) %>% 
  pivot_wider(
    names_from = "member_casual",
    values_from = "number_of_rides"
  ) %>% 
  arrange(year, month)

member_table_df
```
The highest number of both casual and member bikers was in July 2024, and these numbers seem to increase significantly from January 2024. From January 2020 to January 2024, the number of casual bikers appears to be consistently significantly lower than the number of member bikers.

## Table 2
```{r}
top_rides_df = group_by(citibike_df, year, month, start_station_name) %>% 
  summarize(number_of_rides = n()) %>% 
  mutate(ride_ranking = min_rank(number_of_rides)) %>% 
  filter(year == "2024", month == "July", min_rank(desc(number_of_rides)) < 6) %>% 
  select(start_station_name, number_of_rides) %>% 
  arrange(desc(number_of_rides))

top_rides_df
```

## Plot 1
```{r}
week = group_by(citibike_df, weekdays, duration) %>% 
  mutate(duration = median(duration)) %>%
  arrange(weekdays) %>% 
  ggplot(aes(x = weekdays, y = duration))+
  geom_boxplot()

week

month = group_by(citibike_df, month, duration) %>% 
  mutate(duration = median(duration)) %>% 
  ggplot(aes(x = month, y = duration))+
  geom_boxplot()

month

year = group_by(citibike_df, year, duration) %>% 
  mutate(duration = median(duration)) %>% 
  ggplot(aes(x = year, y = duration))+
  geom_boxplot()

year
```

The median duration of cycling time appears to be slightly higher on the weekends.The median duration by month appears to be much higher in July than January. The median duration by year is higher in 2020 than in 2024.


## Plot 2

```{r}
by_month = group_by(citibike_df, year, month, rideable_type, duration) %>% 
  filter(year == "2024") %>% 
  ggplot(aes(x = month, y = duration, fill = rideable_type))+
  geom_boxplot()

by_month

by_membership = group_by(citibike_df, year, member_casual, rideable_type, duration) %>% 
  filter(year == "2024") %>% 
  ggplot(aes(x = member_casual, y = duration, fill = rideable_type))+
  geom_boxplot()

by_membership
  
```
The ride duration by month was highest with electric bikes in July. However, there does not appear to be a noticeable difference between the ride durations for the classic and electric bikes in January. Ride duration was highest for casual bikers who used a classic bike, and there was not a significant difference among member bikers between using a classic or electric bike. 

